---
title: "project"
author: "N Selva Kumar"
output: html_document
---

**Synopsis**

In this project we will be building a prediction model to know how a certain user is lifting weights based on data obtained from an accelerometer.

The dataset consists of 5 classes from 6 participants:

. The subject is lifting weights exactly according to the specification (Class A).

. Throwing the elbow to the front (Class B).

. Lifting the dumbbell only halfway (Class C).

. Lowering the dumbbell only halfway (Class D).

. Throwing the hips to the front (Class E).

For more information and description about the dataset, see the official website: <http://groupware.les.inf.puc-rio.br/har>

**Getting the data**

The file "pml-training" will served as our training set.
The file "pml-testing" is a data set without the classes, in other word, we will predict the classes based on our model.

```{r Getting data}

trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"

if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}

dataset <- read.csv("./data/pml-training.csv", na.strings = c("NA", ""))
validation <- read.csv("./data/pml-testing.csv")
```

**Data preprocessing**

Import the necessary packages.
```{r lib}
library(caret)
library(randomForest)
```

Set a seed for reproducibility.

```{r seed}
set.seed(17)
```

Create the data partitions. 70% of it will go to the training set and the rest will be the test set.

```{r, data partition}
inTrain = createDataPartition(y=dataset$classe, p=0.7, list=FALSE)
training = dataset[inTrain,]
testing = dataset[-inTrain,]
```

Eliminate the NA entries.

```{r training remove NA values}

# Vector of all the columns and the number of NA entries
naColumns = sapply(training, function(x) {sum(is.na(x))}) 

# Columns that has NA values
columnsWithNA = names(naColumns[naColumns > 0])

# Remove columns from the training set
training = training[, !names(training) %in% columnsWithNA] 

#Remove unnecessary columns (the first 7 columns)
training <- training[, -c(1:7)]
names(training)
```

Do the same for the validation set

```{rvalidation remove NA values}

# Make a vector of all the columns and the number of NA entries
naColumns = sapply(validation, function(x) {sum(is.na(x))}) 

# Vector with all the columns that has NA values
columnsWithNA = names(naColumns[naColumns > 0]) 

# Remove columns from the validation set.
validation = validation[, !names(validation) %in% columnsWithNA] 

validation <- validation[, -c(1:7)]
names(validation)
```

Do the same for the testing set.

```{r testing remove NA values}
# Make a vector of all the columns and the number of NA entries
naColumns = sapply(testing, function(x) {sum(is.na(x))}) 

# Vector with all the columns that has NA values
columnsWithNA = names(naColumns[naColumns > 0]) 

#Remove columns from the testing set.
testing = testing[, !names(testing) %in% columnsWithNA] 

testing <- testing[, -c(1:7)]
names(testing)
```

Now, we build the prediction model using Random Forest.

```{r accuracy}

testing$classe <- as.factor(testing$classe)
training$classe <- as.factor(training$classe)

model <- randomForest(classe ~ .,   data=training, ntree = 50)
predictions <- predict(model, testing)
confusionMatrix(predictions, testing$classe)
modelAcc <- confusionMatrix(predictions, testing$classe)$overall[[1]]
```

Our model is `r modelAcc` accurate.

Now, we will predict the unknown classes of the validation set.

```{r, valid}
predictions <- predict(model, validation)
predictions
```